---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Function created and called to unzip data from 'activity.zip' file and to read data in DataFrame. Necessay libraries are loaded.

```{r setup}
library(dplyr)
library(ggplot2)
library(scales)

readActivityData <- function() {
  unzip("activity.zip")
  activity <- read.csv("activity.csv")
  activity$date <- as.Date( as.character( activity$date) )
  activity
}
activity = readActivityData()
```



## What is mean total number of steps taken per day?
```{r}
makeBarGraph <- function(activity) {
  activity %>% group_by( date ) %>% summarize( totalSteps = sum(steps))
  byday <- activity %>% group_by( date ) %>% summarize( totalSteps = sum(steps))
  byday$date <- as.Date( as.character(byday$date) )

  g <- ggplot( byday, aes( x = date, y = totalSteps ) ) + geom_bar( stat = 'identity') + scale_x_date( labels = date_format("%m/%d"), breaks = date_breaks("1 week") )
  list(g, byday)
}
meanMed <- function( activity ) {
  list( mean( activity$steps, na.rm = T ), median( activity$steps, na.rm = T) )
}
makeBarGraph( activity )
meanActivity <- meanMed( activity )
```
The mean number of steps is `r meanActivity[[1]]`.  The median number of steps is `r meanActivity[[2]]`.

## What is the average daily activity pattern?

```{r Number4}
intervalTimeSeries <- function( activity ) {
  byInterval <- activity %>% group_by( interval ) %>% summarize( totalSteps = mean(steps, na.rm = T) )
  ggplot( byInterval, aes( x = interval, y = totalSteps ) ) + geom_point() + geom_line()
}
intervalTimeSeries( activity )
```
```{r MaximumSteps}
maxStepsByInterval <- function( activity ) {
  byInterval <- activity %>% group_by( interval ) %>% summarize( totalSteps = mean(steps, na.rm = T) )
  byInterval[ byInterval$totalSteps == max(byInterval$totalSteps),]
}
maxInterval <- maxStepsByInterval( activity )
```


The interval with the maximum number of steps is `r maxInterval$totalSteps`


## Imputing missing values
```{r NumberOfNAs}
numberOfNAs <- function( activity ) {
  activity$countindex = 1
  sum( activity[ is.na(activity$steps), ]$countindex )
}
```
Using the average steps to fill in missing values.
```{r FillMissingSteps}
fillMissingSteps <- function( activity ) {
  mm <- meanMed( activity )
  activity$steps <- ifelse( is.na(activity$steps), mm[[1]], activity$steps)
  activity
}
```
```{r}
  nas <- numberOfNAs( activity )
  makeBarGraph( fillMissingSteps( activity ) )
  meanActivity <- meanMed( fillMissingSteps( activity ) )
```
The number of NAs is `r nas`.  The mean number of steps is `r meanActivity[[1]]`.  The median number of steps is `r meanActivity[[2]]`.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
  applyWeekDay <- function( activity ) {
    activity$weekType <- weekdays( activity$date )
    activity$weekType <- ifelse (activity$weekType %in% c('Saturday', 'Sunday'), 'weekend', 'weekday')
    activity$weekType <- factor( activity$weekType)
    activity
  }

  plotByWeekType <- function( activity ) {
    ggplot( activity, aes( x = interval, y = steps, group = weekType, color = weekType)) + geom_line() + facet_wrap( ~ weekType)
  }
  activityFilled <- fillMissingSteps( activity)
  activityWeekType <- applyWeekDay( activityFilled )
  head( activityWeekType )
  plotByWeekType( activityWeekType )
```

Yes there are differences.  More steps are taken on the weekdays.